
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)

# Load the dataset
file_path <- "C:/Users/manoh/Downloads/Compressed/archive_2/final_data.csv"
data <- read_csv(file_path)
print(data)
summary(data)
```

Question 1: What is the Distribution of Points Scored by Players?
To visualize the distribution of points scored by players, we can use a histogram.
```{r}
library(ggplot2)

# Plot histogram of points scored
ggplot(data, aes(x = PTS)) +
  geom_histogram(binwidth = 5, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Points Scored by Players", x = "Points Scored", y = "Frequency") +
  theme_minimal()

```

Question 2: How Do Field Goals Made (FGM) and Three-Point Field Goals Made (FG3M) Relate to Each Other?
To visualize the relationship between field goals made and three-point field goals made, we can use a scatter plot.
```{r}
# Plot scatter plot of FGM vs. FG3M
ggplot(data, aes(x = FGM, y = FG3M)) +
  geom_point(color = "blue", alpha = 0.6) +
  labs(title = "Relationship Between Field Goals Made and Three-Point Field Goals Made", x = "Field Goals Made (FGM)", y = "Three-Point Field Goals Made (FG3M)") +
  theme_minimal()

```
Question 3: What Are the Average Field Goals Made (FGM), Blocks (BLK_y), and Rebounds (REB) by Team?
To visualize the average field goals made, blocks, and rebounds by team, we can use a bar plot.
```{r}
# Calculate the average FGM, BLK_y, and REB by team
team_averages <- data %>%
  group_by(TEAM_NAME_y) %>%
  summarise(
    Avg_FGM = mean(FGM, na.rm = TRUE),
    Avg_BLK_y = mean(BLK_y, na.rm = TRUE),
    Avg_REB = mean(REB, na.rm = TRUE)
  ) %>%
  gather(key = "Statistic", value = "Average", -TEAM_NAME_y)

# Plot bar plot of average FGM, BLK_y, and REB by team
ggplot(team_averages, aes(x = TEAM_NAME_y, y = Average, fill = Statistic)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Average Field Goals Made, Blocks, and Rebounds by Team", x = "Team", y = "Average Value") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Question 4: What is the Distribution of Player Efficiency Ratings?
To visualize the distribution of player efficiency ratings, we can use a histogram.
```{r}
# Plot histogram of player efficiency ratings
ggplot(data, aes(x = Efficiency)) +
  geom_histogram(binwidth = 5, fill = "green", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Player Efficiency Ratings", x = "Efficiency Rating", y = "Frequency") +
  theme_minimal()

```

Question 5: How Does the Number of Wins Compare Between Home and Away Games?
To compare the number of wins between home and away games, we can use a grouped bar plot.
```{r}
# Summarize the number of wins by location
wins_by_location <- data %>%
  group_by(LOCATION) %>%
  summarise(Total_Wins = sum(W_x, na.rm = TRUE))

# Plot grouped bar plot of wins by location
ggplot(wins_by_location, aes(x = LOCATION, y = Total_Wins, fill = LOCATION)) +
  geom_bar(stat = "identity") +
  labs(title = "Comparison of Wins in Home and Away Games", x = "Location", y = "Total Wins") +
  theme_minimal()

```
Question 6: What is the Average Plus/Minus Statistic by Player Position?
To visualize the average plus/minus statistic by player position, we can use a box plot.
```{r}
# Calculate the average plus/minus statistic by player position
plus_minus_by_position <- data %>%
  group_by(POSITION) %>%
  summarise(Average_Plus_Minus = mean(PLUS_MINUS, na.rm = TRUE))

# Plot box plot of average plus/minus by player position
ggplot(data, aes(x = POSITION, y = PLUS_MINUS)) +
  geom_boxplot(fill = "orange", color = "black") +
  labs(title = "Average Plus/Minus Statistic by Player Position", x = "Player Position", y = "Plus/Minus") +
  theme_minimal()

```

Question 7: How Do Free Throws Made (FTM) Vary Across Teams?
To visualize the variation of free throws made across different teams, we can use a violin plot
```{r}
# Plot violin plot of free throws made across teams
ggplot(data, aes(x = TEAM_NAME_y, y = FTM)) +
  geom_violin(fill = "purple", alpha = 0.7) +
  labs(title = "Variation of Free Throws Made Across Teams", x = "Team", y = "Free Throws Made") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Question 8: What is the Correlation Matrix of Key Statistics?
To visualize the correlation matrix of key statistics (e.g., FGM, FG3M, REB, AST, PTS), we can use a correlation heatmap.
```{r}
library(tidyr)
library(ggplot2)
library(tibble)

# Select key statistics
key_stats <- data %>%
  select(FGM, FG3M, REB, AST, PTS)

# Calculate the correlation matrix
correlation_matrix <- cor(key_stats, use = "complete.obs")

# Convert correlation matrix to a tidy format
melted_correlation_matrix <- as.data.frame(correlation_matrix) %>%
  rownames_to_column(var = "Var1") %>%
  pivot_longer(cols = -Var1, names_to = "Var2", values_to = "value")

# Plot heatmap of correlation matrix
ggplot(data = melted_correlation_matrix, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Correlation") +
  theme_minimal() +
  labs(title = "Correlation Matrix of Key Statistics", x = "", y = "") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, 
                                   size = 12, hjust = 1))

```
Question 9: What is the Relationship Between Turnovers (TOV) and Steals (STL_x)?
To visualize the relationship between turnovers and steals, we can use a bubble plot where the size of the points represents the number of games played (GP_x).
```{r}
# Plot bubble plot of turnovers vs. steals
ggplot(data, aes(x = TOV, y = STL_x, size = GP_x, color = TEAM_NAME_y)) +
  geom_point(alpha = 0.6) +
  labs(title = "Relationship Between Turnovers and Steals", x = "Turnovers", y = "Steals", size = "Games Played") +
  theme_minimal() +
  theme(legend.position = "none")

```


Explanation of Graphs:
Histogram of Points Scored: This graph helps visualize the frequency distribution of points scored by players, showing how many players fall into various point ranges.

Scatter Plot of FGM vs. FG3M: This graph illustrates the relationship between the number of field goals made and the number of three-point field goals made, helping to identify any correlations.

Bar Plot of Average FGM, BLK_y, and REB by Team: This graph shows the average field goals made, blocks, and rebounds for each team, allowing for comparison across teams.

These visualizations will help in understanding the data better and drawing meaningful insights. Let me know if you need further adjustments or additional questions!
